---
title: "The effect of Cyclones on Climate Awareness and Happiness (WIP)"
author: "Group 12"
date: "November 12th, 2024"
output: pdf_document
---

List your group members, including their student numbers, here:

- Joshua Goudy (169031329)
- Melissa Thambiaiah (169060509)
- Emaan Umar (169108097)
- Joseph Egbo (143037470)

Ideas: Cyclones, climate change, and happiness

- Does a cyclone hitting a country increase climate awareness and decrease happiness

```{r setup, include=FALSE}
# echo = FALSE will set the Rmd to *not* show the R code. Don't change this.
# You may change the default figure width and figure height as you please.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 6)

# Put any libraries that you need to load here.
# DO NOT PUT "install.packages()" IN AN RMD FILE!!!
library(tidyverse)
library(arrow)
```

Note that all of the code is *hidden* by default. This file will be graded based on the insights, not the code.

You will only submit the PDF version of this document. To knit to PDF, you'll need to run `install.packages("tinytex")` in the console, followed by `tinytex::install_tinytex()` (DO NOT PUT THESE COMMANDS IN AN RMD FILE!!!). If you encounter errors in "Knit to PDF", you can "knit to html" and then print the html file to PDF using your operating system's PDF view (e.g. Adobe Acrobat). Only standalone PDF files will be accepted by MyLS.

# Abstract

Ideas: How does a cyclone hitting a country affect the climate awareness and overall happiness.


# Introduction

In this report, we are going to delve into the correlation between cyclones, climate awareness and the way in which happiness levels fluctuate based on the occurrences of cyclones. To establish a sense of importance on these topics, let's dive into the history of Hurricane Katrina. Making its mark in American history, there is empirical evidence that reveals how Hurricane Katrina has dictated happiness levels, while promoting the significance of climate awareness. 
         
Moreover, we are applying exploratory techniques by visualizing the graphical representations of these data sets in order to conduct an effective comparison between the information sets. 

Ultimately, our findings will depict how cyclones/hurricanes will play a role in determining happiness levels and the rates of climate awareness, which will convey the analysis of causation.


# Data Description

## Cyclones Data

```{r load_data1}
cyclone_data <- read_parquet('cyclones_data.parquet')
```

The data comes from NOAA Cyclone/Hurricane Data contains record of tropical storms over the past 173 years.

In order to clean/tidy the data, we separated each different type of data entry into a unique column, filtered out irrelevant data entries, converted the location data to a readable format, and classified the storms.


## Climate Awareness Data

```{r load_data2}
climate_awareness_data <- read_parquet('climate_awareness.parquet')
```

The data come from 2022 Climate Change Opinion Survey and detail climate change opinion records how aware the population of various countries are about climate change, going forward this data set will be referred to as 'climate awareness'.

In order to clean the data, we type of data entry into a unique column, filtered out irrelevant data entries, and removed outdated entries.

## Happiness Data

```{r load_data3}
happiness_data <- read_parquet('happiness.parquet')
```

The data come from World Happiness Report Score (from 2023 report) and records various statistics about countries that are believed to impact the overall happiness of a country and gives them an overall happiness score, going forward this data set will be referred to as 'happiness'.

In order to clean/tidy the data, we separated each different removed repeated data, filtered out countries without a lifeladder score, and filtered out the outdated entries.


## Combining the Data

Explain how any combinations of data were performed. Explain what kind of join was needed, whether columns had to be modified (for example, matching "country" names.)


# Exploratory Data Analysis

To achieve our goals, we explored the data by...

We explored many aspects of the data, but will demonstrate three. These are the correlation between happiness & cyclones/hurricanes hitting a country, the correlation between cyclones/hurricanes hitting a country and how aware its people are of climate change, and the correlation between how aware the people of a country are and how happy those people are.

The first aspect that we found interesting is shown in \@ref(fig:insight1). The insight should be specific to the data shown, not a general statement beyond the data (leave that for the conclusion).

```{r insight1, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
# This is an example of how you can control figures and captions in
# an R chunk. Note that you can reference figures using:
# \@ref(fig:insight1), where "insight1" is the label of this code
# chunk (the first bit of text after the "r" in "```{r label, options...}")
```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

```{r summary_stats}
# Calculate the relevant summary statistics here.
# Note that the "kable" function in the "knitr" package
# is convenient for making nice tables. Other packages can
# do much fancier things with tables, but keep in mind that
# the insights should be the star, not the formatting.
```

The next insight that we found is shown in \@ref(fig:insight2).

```{r insight2, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
# This figure will have a height of 4 and a width of 6.
# Feel free to change this, and to apply different sizes
# to the other figures you create.
```

Finally, \@ref(fig:insight3) shows ...

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
```

# Conclusion and Future Work

Overall, we found <<general ideas>>.

A second paragraph about our findings.

The next steps in this analysis are...

The limitations of this analysis are as follows. (Do not simply list potential issues with sampling, but relate them to your analysis and how they affect your conclusions. An honest and complete acknowledgement of the limitations makes the analysis more trustworthy.)

# References

I am not strict about MLA or APA style or anything like that. For this report, I would much rather have your citations be easy to match to your insights.

The easiest way is to use Rmd's [footnote](https://bookdown.org/yihui/rmarkdown/markdown-syntax.html#inline-formatting) syntax. This will put a number beside the word where the footnote appears, and the full text of the footnote at the bottom of the page (pdf) or end of the document (html). The syntax is:^[See the source view to see this footnote], where I suggest that you put in something like this^[The relevance to the insight is ... . From <<name of source and name of article>>, published on <<date>>, url: <<link to page>>] to make references for this assignment.

Alternatively, you could make a list of citations with their main arguments and why they're relevent to your insights, methods, etc.

The link above also references "bibtex" files. These are also extremely convenient, but have a steep learning curve and they make it difficult to tie them to an insight. If you use bibtext, then make sure that you provide a sentence to describe the source and it's relevance when you cite it - don't just add citations to the end of a sentence (this is common practice in academia, but I want to know that your citations are directly relevant for this assignmnet).
